---
title: "Biostat 203B Homework 1"
author: Alvin Wang
subtitle: Due Jan 22 @ 11:59PM
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Display machine information for reproducibility:
```{r, eval=F}
sessionInfo()
```

## Q1. Git/GitHub

**No handwritten homework reports are accepted for this course.** We work with Git and GitHub. Efficient and abundant use of Git, e.g., frequent and well-documented commits, is an important criterion for grading your homework.

1. Apply for the [Student Developer Pack](https://education.github.com/pack) at GitHub using your UCLA email.

2. Create a **private** repository `biostat-203b-2021-winter` and add `Hua-Zhou`, `Chris-German` and `ElvisCuiHan` as your collaborators with write permission.

3. Top directories of the repository should be `hw1`, `hw2`, ... Maintain two branches `master` and `develop`. The `develop` branch will be your main playground, the place where you develop solution (code) to homework problems and write up report. The `master` branch will be your presentation area. Submit your homework files (R markdown file `Rmd`, `html` file converted from R markdown, all code and data sets to reproduce results) in `master` branch.

4. After each homework due date, teaching assistant and instructor will check out your master branch for grading. Tag each of your homework submissions with tag names `hw1`, `hw2`, ... Tagging time will be used as your submission time. That means if you tag your `hw1` submission after deadline, penalty points will be deducted for late submission.

5. After this course, you can make this repository public and use it to demonstrate your skill sets on job market.

## Q2. Linux Shell Commands

1. This exercise (and later in this course) uses the [MIMIC-IV data](https://mimic-iv.mit.edu), a freely accessible critical care database developed by the MIT Lab for Computational Physiology. Follow the instructions at <https://mimic-iv.mit.edu/docs/access/> to (1) complete the CITI `Data or Specimens Only Research` course and (2) obtain the PhysioNet credential for using the MIMIC-IV data. Display the verification links to your completion report and completion certificate here. (Hint: The CITI training takes a couple hours and the PhysioNet credentialing takes a couple days; do not leave it to the last minute.)  
[CITI completion report](https://www.citiprogram.org/verify/?kb31c0f6e-150c-4078-bb91-4ebf632e8e20-40374259)  
[CITI completion certificate](https://www.citiprogram.org/verify/?w4e9cfcc0-217a-4601-a4a1-882c6077063f-40374259)

2. The `/usr/203b-data/mimic-iv/` folder on teaching server contains data sets from MIMIC-IV. Refer to <https://mimic-iv.mit.edu/docs/datasets/> for details of data files.  
    ```{bash}
    ls -l /usr/203b-data/mimic-iv
    ```
Please, do **not** put these data files into Git; they are big. Do **not** copy them into your directory. Do **not** decompress the gz data files. These create unnecessary big files on storage and are not big data friendly practices. Just read from the data folder `/usr/203b-data/mimic-iv` directly in following exercises. 

    Use Bash commands to answer following questions.

3. Display the contents in the folders `core`, `hosp`, `icu`. What are the functionalities of the bash commands `zcat`, `zless`, `zmore`, and `zgrep`?  
```{bash}
ls -l /usr/203b-data/mimic-iv/*
```

`zcat` is  for viewing the contents of a compressed file without literally uncompressing it.  
`zless` is a filter which allows examination of compressed or plain text files one screenful at a time.  
`zmore` command pauses after each screenful, printing --More-- at the bottom of the screen.  
`zgrep` command is used to search out expressions from a given a file even if it is compressed. 

4. What's the output of following bash script?
    ```{bash}
    for datafile in /usr/203b-data/mimic-iv/core/*.gz
      do
        ls -l $datafile
      done
    ```
Display the number of lines in each data file using a similar loop.
    ```{bash}
    for datafile in /usr/203b-data/mimic-iv/core/*.gz
      do
        wc -l $datafile
      done
    ```

4. Display the first few lines of `admissions.csv.gz`. How many rows are in this data file? How many unique patients (identified by `subject_id`) are in this data file? What are the possible values taken by each of the variable `admission_type`, `admission_location`, `insurance`, `language`, `marital_status`, and `ethnicity`? Also report the count for each unique value of these variables. (Hint: combine Linux commands `zcat`, `head`/`tail`, `awk`, `uniq`, `wc`, and so on.)
```{bash}
zcat /usr/203b-data/mimic-iv/core/admissions.csv.gz | head
```

```{bash}
#rows count
tail -n +2 /usr/203b-data/mimic-iv/core/admissions.csv.gz | wc -l
```

```{bash}
#number of unique patients
zcat /usr/203b-data/mimic-iv/core/admissions.csv.gz | awk -F ',' '{ print $1 }' | sort | uniq | wc -l
```

```{bash}
#possible values and their counts for `admission_type`
zcat /usr/203b-data/mimic-iv/core/admissions.csv.gz | tail -n +2 | awk -F ',' '{ print $6 }' | sort | uniq -c
```

```{bash}
#possible values and their counts for `admission_location`
zcat /usr/203b-data/mimic-iv/core/admissions.csv.gz | tail -n +2 | awk -F ',' '{ print $7 }' | sort | uniq -c
```

```{bash}
#possible values and their counts for `insurance`
zcat /usr/203b-data/mimic-iv/core/admissions.csv.gz | tail -n +2 | awk -F ',' '{ print $9 }' | sort | uniq -c
```

```{bash}
#possible values and their counts for `language`
zcat /usr/203b-data/mimic-iv/core/admissions.csv.gz | tail -n +2 | awk -F ',' '{ print $10 }' | sort | uniq -c
```

```{bash}
#possible values and their counts for `marital_status`
zcat /usr/203b-data/mimic-iv/core/admissions.csv.gz | tail -n +2 | awk -F ',' '{ print $11 }' | sort | uniq -c
```

```{bash}
#possible values and their counts for `ethnicity`
zcat /usr/203b-data/mimic-iv/core/admissions.csv.gz | tail -n +2 | awk -F ',' '{ print $12 }' | sort | uniq -c
```




## Q3. Who's popular in Price and Prejudice

1. You and your friend just have finished reading *Pride and Prejudice* by Jane Austen. Among the four main characters in the book, Elizabeth, Jane, Lydia, and Darcy, your friend thinks that Darcy was the most mentioned. You, however, are certain it was Elizabeth. Obtain the full text of the novel from <http://www.gutenberg.org/cache/epub/42671/pg42671.txt> and save to your local folder. 
    ```{bash}
    curl http://www.gutenberg.org/cache/epub/42671/pg42671.txt > pride_and_prejudice.txt
    ```
Do **not** put this text file `pride_and_prejudice.txt` in Git. Using a `for` loop, how would you tabulate the number of times each of the four characters is mentioned?
```{bash}
  for char in {'Elizabeth','Jane','Lydia','Darcy'}
    do
      grep -o $char pride_and_prejudice.txt | wc -l
    done
```


2. What's the difference between the following two commands?
    ```{bash eval=FALSE}
    echo 'hello, world' > test1.txt
    ```
    and
    ```{bash eval=FALSE}
    echo 'hello, world' >> test2.txt
    ```
`>` writes 'hello, world' to test1.txt.  
`>>` appends 'hello, world' to test2.txt.


3. Using your favorite text editor (e.g., `vi`), type the following and save the file as `middle.sh`:
    ```{bash eval=FALSE}
    #!/bin/sh
    # Select lines from the middle of a file.
    # Usage: bash middle.sh filename end_line num_lines
    head -n "$2" "$1" | tail -n "$3"
    ```
Using `chmod` make the file executable by the owner, and run 
    ```{bash eval=FALSE}
    ./middle.sh pride_and_prejudice.txt 20 5
    ```
Explain the output. Explain the meaning of `"$1"`, `"$2"`, and `"$3"` in this shell script. Why do we need the first line of the shell script?
```{bash eval=FALSE}

Author: Jane Austen

Editor: R. W. (Robert William) Chapman

```

Output is the line 16-20 of pride_and_prejudice.txt. (Last 5 lines of the first 20 lines.)  
$1 takes the first argument, which is the file name here.  
$2 takes the second argument, which is the number of end line. (head -n)  
$3 takes the third argument, which is the number of last lines. (tail)  

### Q4. More fun with Linux

Try these commands in Bash and interpret the results: `cal`, `cal 2021`, `cal 9 1752` (anything unusual?), `date`, `hostname`, `arch`, `uname -a`, `uptime`, `who am i`, `who`, `w`, `id`, `last | head`, `echo {con,pre}{sent,fer}{s,ed}`, `time sleep 5`, `history | tail`.
`cal` prints the calender of current month.  
```{bash}
cal
```
`cal 2021` prints the calender of the whole year. 
```{bash}
cal 2021
```
`cal 9 1752` prints the calender of September 1752. Notice today is not marked because it was in the past.  
```{bash}
cal 9 1752
```
`date` prints the current date and time.  
```{bash}
date
```
`hostname` prints the current host name.  
```{bash}
hostname
```
`arch` prints the computer architecture. 
```{bash}
arch
```
`uname -a` prints system information and information about the machine and operating system it is run on.  
```{bash}
uname -a
```
`uptime` dsplays the time since the last boot.  
```{bash}
uptime
```
`who am i` displays current user. 
```{bash}
who am i
```
`who` print who is currently logged in.  
```{bash}
who
```
`w` displays information about currently logged in users and what each user is doing.  
```{bash}
w
```
`id` displays information about user id. 
```{bash}
id
```
`last | head` displays user logging history.
```{bash}
last | head
```
`echo {con,pre}{sent,fer}{s,ed}` prints all the permutations of the the three pairs.  
```{bash}
echo {con,pre}{sent,fer}{s,ed}
```
`time sleep 5` pauses execution for 5 seconds. 
```{bash}
time sleep 5
```
`history | tail` reports the last few commands.  
```{bash}
history | tail
```